---
title: "Rerewhakaaitu"
author: "Abby Lewis"
date: "2024-03-11"
output: html_document
---

For continuous data, the output format has three columns (with one year of data):

-   `yday`: day of year (numeric)
-   `Variable`: Variable name
-   `Value`: Variable value

For discrete data, the output format has four columns

-   `start`: start date of this "season" as numeric day of year
-   `end`: end date of this season as numeric day of year
-   `Season`: season name (e.g., "summer", "stratified")
-   `Method`: method name (i.e., y-axis label in figure 3)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rLakeAnalyzer)

offset = as.Date("2023-01-01") - as.Date("2022-06-21")
LAKE = "Lake Rerewhakaaitu (2022-2023)" 
```

In-lake data (temperature, chlorophyll-a, stratification)

```{r}
# Load data
in_lake_raw <- read_csv("https://pasta-s.lternet.edu/package/data/eml/edi/1312/2/f8e0e7b22826772424f156b9565d1498") %>%
  filter(Lake == "Lake Rerewhakaaitu") %>%
  mutate(day = as.Date(Date) + offset,
         yday = yday(day))
in_lake_data <- in_lake_raw %>%
  filter(year(day) == 2023,
         !is.na(yday)) 

# Format continuous data
in_lake_formatted <- in_lake_data %>%
  select(-Chla_ugL) %>%
  pivot_longer(c("WaterTemperature_C", "DO_mgL", "Chla_fluorescence"),
               values_to = "Value") %>%
  mutate(Variable = paste(Layer, name),
         Variable = case_match(Variable,
                               "Surface WaterTemperature_C"~"Surface_temperature",
                               "Surface DO_mgL"~"Surface_DO",
                               "Bottom WaterTemperature_C"~"Bottom_temperature",
                               "Bottom DO_mgL"~"Bottom_DO",
                               "Surface Chla_fluorescence"~"Surface_chla")) %>%
  filter(!is.na(Variable)) %>%
  mutate(Lake = LAKE) %>%
  select(Variable, Value, yday, Lake)

write.csv(in_lake_formatted, "../01b_Processed_data/Rerewhakaaitu_continuous.csv", row.names = F)

# Stratification metric calcs (both discrete and continuous)
strat_export_cont <- in_lake_data %>%
  select(yday, WaterTemperature_C, Layer) %>%
  pivot_wider(names_from = Layer, values_from = WaterTemperature_C) %>%
  group_by(yday) %>%
  summarize(Diff_Dens_surf_bot = water.density(Bottom) - water.density(Surface)) %>%
  pivot_longer(-1, names_to = "Variable", values_to = "Value") %>%
  mutate(Lake = LAKE)

strat_export_discrete <- strat_export_cont %>%
  mutate(Strat = ifelse(Value > 0.1, "Stratified", "Mixed")) %>%
  filter(!is.na(Value)) %>%
  ungroup() %>%
  filter(Strat != lag(Strat)) %>%
  full_join(data.frame(Variable = "Diff_Dens_surf_bot", yday = 1, Strat = "Mixed")) %>%
  arrange(yday) %>%
  mutate(end = lead(yday, default = 365)) %>%
  rename(start = yday, Season = Strat) %>%
  select(start, end, Season) %>%
  mutate(Lake = LAKE)

write.csv(strat_export_discrete, 
          "../01b_Processed_data/Rerewhakaaitu_stratification_discrete.csv", 
          row.names = F)
write.csv(strat_export_cont, 
          "../01b_Processed_data/Rerewhakaaitu_stratification_continuous.csv", 
          row.names = F)
```

Meteorological data

```{r}
# Load data
met <- read_csv("https://pasta-s.lternet.edu/package/data/eml/edi/1312/2/e312838eff13f56594021498f35c04cd")

# Format continuous data
met_formatted <- met %>%
  filter(Lake == "Lake Rerewhakaaitu") %>%
  mutate(Date = as.Date(Date) + offset,
         doy = yday(Date)) %>%
  filter(year(Date) == 2023) %>%
  rename(AirTemp_C_Average = AirTemp_C,
         Shortwave = ShortwaveRadiationClear_Wm2) %>%
  pivot_longer(c(AirTemp_C_Average, Shortwave), names_to = "Variable", values_to = "Value") %>%
  mutate(yday = yday(Date)) %>%
  select(Variable, Value, yday) %>%
  mutate(Lake = LAKE)

write.csv(met_formatted, "../01b_Processed_data/Rerewhakaaitu_met.csv", row.names = F)
```

Ice data
(no ice)
```{r}
ice <- data.frame(Lake = LAKE,
                  Season = "Open water",
                  Method = "Ice",
                  start = 0,
                  end = 365)
write.csv(ice, "../01b_Processed_data/Rerewhakaaitu_ice.csv", row.names = F)
```

