---
title: "Template"
author: "Abby Lewis"
date: "2024-03-11"
output: html_document
---

# Rerewhakaaitu

For continuous data, the output format has three columns (with one year of data):

-   `yday`: day of year (numeric)
-   `Variable`: Variable name
-   `Value`: Variable value

For discrete data, the output format has four columns

-   `start`: start date of this "season" as numeric day of year
-   `end`: end date of this season as numeric day of year
-   `Season`: season name (e.g., "summer", "stratified")
-   `Method`: method name (i.e., y-axis label in figure 3)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rLakeAnalyzer)
```

In-lake data (temperature, chlorophyll-a, stratification)

```{r}
# Load data
in_lake_raw <- read.csv("../01a_Raw_data/Rerewhakaaitu_buoy_wq.csv")
in_lake_data <- in_lake_data %>%
  rename(Value = avg_value,
         Variable = variable,
         yday = doy) %>%
  filter(year == 2023) 

# Format continuous data
in_lake_formatted <- in_lake_data %>%
  filter(depth_rnd == 1) %>%
  select(Variable, Value, yday) %>%
  mutate(Variable = ifelse(Variable == "TmpWtr", "Surface_temperature", Variable),
         Variable = ifelse(Variable == "FlChlr", "Surface_chla", Variable)) %>%
  mutate(Lake = "Rerewhakaaitu")

write.csv(in_lake_formatted, "../01b_Processed_data/Rerewhakaaitu_continuous.csv", row.names = F)

# Stratification metric calcs (both discrete and continuous)
strat_export_cont <- in_lake_data %>%
  filter(Variable == "TmpWtr") %>%
  pivot_wider(names_from = depth_rnd, values_from = Value, names_prefix = "Temp_") %>%
  group_by(yday) %>%
  summarize(Diff_Dens_surf_bot = water.density(Temp_12) - water.density(Temp_1)) %>%
  pivot_longer(-1, names_to = "Variable", values_to = "Value") %>%
  mutate(Lake = "Rerewhakaaitu")

strat_export_discrete <- strat_export_cont %>%
  mutate(Strat = ifelse(Value > 0.1, "Stratified", "Mixed")) %>%
  filter(!is.na(Value)) %>%
  filter(!Strat == lag(Strat)) %>%
  full_join(data.frame(Variable = "Diff_Dens_surf_9", yday = 0, Strat = "Stratified")) %>%
  arrange(yday) %>%
  mutate(end = lead(yday, default = 365)) %>%
  rename(start = yday, Season = Strat) %>%
  filter(Season == "Stratified") %>%
  select(start, end, Season) %>%
  mutate(Lake = "Rerewhakaaitu")

write.csv(strat_export_discrete, 
          "../01b_Processed_data/Rerewhakaaitu_stratification_discrete.csv", 
          row.names = F)
write.csv(strat_export_cont, 
          "../01b_Processed_data/Rerewhakaaitu_stratification_continuous.csv", 
          row.names = F)
```

Meteorological data

```{r}
# Load data
met <- read.csv("../01a_Raw_data/Rerewhakaaitu_meteorology.csv")

# Format continuous data
met_formatted <- met %>%
  filter(year(DateTime) == 2023) %>%
  mutate(variable = ifelse(variable == "TmpAir", "AirTemp_C_Average", variable),
         variable = ifelse(variable == "RadClr", "Shortwave", variable),
         Date = as.Date(DateTime)) %>%
  filter(variable %in% c("AirTemp_C_Average", "Shortwave")) %>%
  rename(Variable = variable) %>%
  group_by(Date, Variable) %>%
  summarize(Value = mean(value, na.rm = T), .groups = "drop") %>%
  mutate(yday = yday(Date)) %>%
  select(-Date) %>%
  mutate(Lake = "Rerewhakaaitu")

write.csv(met_formatted, "../01b_Processed_data/Rerewhakaaitu_met.csv", row.names = F)
```

Ice data

```{r}
#Load data
ice <- read.csv("../01a_Raw_data/YOUR LAKE ice.csv")

ice_formatted <- ice #apply necessary transformations

write.csv(ice_formatted, "../01b_Processed_data/YOUR LAKE_ice.csv", row.names = F)
```
